---
title: "Reading OneDrive and SharePoint files with read365 R package"
author: "Artur Quirino"
date: "2023-02-08"
date-modified: "2023-02-20"
categories:
  - R
  - read365
---

In this post, I would like to introduce four functions available in the `read365`
package that allow for the reading of OneDrive and SharePoint files: `OneDrive_file`,
`SharePoint_file`, `OneDrive_shared_file` and `SharePoint_shared_file`. 
The former two functions are intended for private content, while the last two are 
designed for public files. It is worth noting that `read365` is heavily reliant on the 
`Microsoft365R` package.

The goal of the `read365` package is to allow users to read files from office365 
directly in R, without the need to sync files between their computer and the cloud.

## Reading private content from OneDrive

The `OneDrive_file` function executes the login process onto OneDrive by 
utilizing Azure Active Directory authentication. Upon successful authentication, 
`OneDrive_file` requires users to set three parameters: the location of the file (`file`),
the desired function for reading the file (`.function`), and optionally, arguments for this function (`...`). 
See the example below.

```{r}
#remotes::install_github("arturhgq/read365")
read365::OneDrive_file(
  "CienciaPolitica/data/read365_examples/OneDrive_file.xlsx",
  readxl::read_excel, 
  sheet = 1
)
```

## Reading private content from SharePoint

Shortly, the usage of the `SharePoint_file` is similiar to `OneDrive_file`. 
However, there are two additional parameters that must be set: `SharePoint` 
and `drive`. In the `SharePoint` parameter, the user is required to specify the 
SharePoint URL where the target file is located. As SharePoint may contain multiple 
drives to store files, users must indicate the specific `drive`[^1] to be accessed too.
In most cases, "Documents" is the primary drive for storing files in SharePoint.

[^1]: In case the appropriate drive is unknown, a list of drives can be obtained by 
utilizing the `read365::get_drives` function.

```{r}
read365::SharePoint_file(
  SharePoint = "https://arturhgq.sharepoint.com",
  drive = "Documentos",
  file = "Sharepoint_file.xlsx",
  .function = readxl::read_excel
)
```

## Reading public content from OneDrive or SharePoint
 
To use `OneDrive_shared_file` or `SharePoint_shared_file`[^2] no authentication is required. Therefore, all that is
needed is to provide the file's URL along with the corresponding function for reading 
the file and, optionally, its respective parameters.

[^2]: Internally, both functions are identical.

```{r}
 url_OneDrive = 'https://arturhgq-my.sharepoint.com/:x:/p/contact/EZ_KJ3cqtIVEh4PdXhTGy7IBpWT5_Zlp2VjYwlgVCPK4oQ?e=UZGQZt'
 url_SharePoint = 'https://arturhgq.sharepoint.com/:x:/g/EZOMb55WRsZFhXuo5K8uGXYBaRdGPNhX7lFNDjLpphj-mw?e=nCm0L4'
 
 read365::OneDrive_shared_file(
    url_OneDrive,
    .function = readxl::read_excel,
    sheet = 1
  )

  read365::SharePoint_shared_file(
    url_SharePoint,
    .function = readxl::read_excel,
    sheet = 1
  )
```





